---
title: "BBkVignette"
author: Daniel Ollech
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BBkVignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Quick intro DSA2
The first version of dsa was developed by Ollech (2018, 2021). Based on the analysis of the daily currency in circulation, it was shown that daily time series can comprise calendar effects and several periodic effects. The latter can be handled by traditional seasonal adjustment methods, such as STL (see Cleveland et al, 1990, for details on this local regressions based method).

While the first version of dsa used a non-seasonal ARIMA model with trigonometric components and additional calendar regressors to conduct the calendar and outlier adjustment, DSA2  uses the extended airline model developed by Jean Palate and included in the {rjd3highfreq} package. This allows the estimation of an Airline type ARIMA model with multiple periodic effects. Therefore, the order of the steps in dsa has been adapted as follows:

1) Extended Airline model to estimate calendar and outlier effects
2) Estimation of the day-of-the-week effect
3) Estimation of the day-of-the-month effect
4) Estimation of the day-of-the-year effect

Steps 2 to 4 may be iterated several times to improve the estimation of the periodic effects. For the estimation of these effects, different methods can be used. Currently, X-11, STL and Seats (see below) are implemented.

## Setup

To get started, ensure that you have Java 17 or higher. If not automatically detected, 
you can set your Java home using the following command.
 
```{r start}
Sys.setenv("JAVA_HOME" = "C:\\Workspace\\Java\\JDK\\jdk-17.0.3+7") ## Always start with this
```

**At the Bundesbank, you can get Java 17 here:** 

**R:/Zentrale/Projekte/FB-S/Daten/DA/JDemetra/Anwendungen/VC/Java/JDK**

**Move and unpack the zip file to the right path: C:/Workspace/Java/JDK/**

## Getting started with dsa2

Once your Java settings are up to speed, you can start using the {dsa2} package
```{r setup}
library(dsa2)
```

## Creating some seasonal time series
For simplicity this example only includes the Easter holidays as holiday regressors, as they are the prime example of calendar effects. 
```{r createSeries}
## Creating series with calendar effects, and a day-of-the-week and day-of-the-year pattern

set.seed(135)
all <- tssim::sim_daily(N = 6, 
                        week_sd = 10,
                        year_sd = 10,
                        extra_smooth = TRUE,
                        calendar = list(which = "EasterMonday", 
                                        effect_size = 40, 
                                        from = -1, 
                                        to = 0))
series <- all$original

## Creating holiday regressors
regs <- holidaysCentered[ ,c("EasterMonday", 
                            "EasterSunday")][seq.Date(from = start(series), 
                                                      length.out = length(series) + 365, 
                                                      by = "days")]

## Creating series with only day-of-the-month pattern
allm <- tssim::sim_daily(N = 5, 
                        week_sd = 0,
                        year_sd = 0,
                        month_sd = 10)

series_withMonthlySeasonality <- allm$original
```



## Default adjustment

Currently, by default, STL is used for the seasonal adjustments and no automatic calendar adjustment is conducted
```{r Default, fig.width=15, fig.height=4.5, fig.align = 'center'}
result0 <- dsa(series)
result0
plot(result0)
```


## Create HTML output

To create an HTML output you can use the following code

```{r output, fig.width=11, fig.height=4.5, eval = FALSE}
output(result0) # Create HTML output
```

## Adjustment with regressors

To conduct a calendar adjustment, you have to include your pre-defined regressors, as follows

```{r DefaultRegs, fig.width=15, fig.height=4.5, fig.align = 'center'}
result <- dsa(series,
              xreg = regs)
result
plot(result)
```


## Adjustment with missing values
dsa() handles missing values automatically, they are filled up using the fractional airline model

```{r AdjustmentMissing, fig.width=15, fig.height=4.5, fig.align = 'center'}
seriesx <- series
seriesx[c(1, 499, 1010, 1748)] <- NA

resultx <- dsa(seriesx,
              xreg = regs)
resultx
plot(resultx)
```




## Adjustment with X-11

The parameters s7, s31, and s365 control for the method or specification used for the adjustment of day-of-the-week, day-of-the-month, and day-of-the-year effects. Setting any of these parameters (s7, s31 or s365) to "x11" uses the default adjustment of x11. Alternatively, x11_method() can be used, to specify details of the adjustment.

The function compare_plot() allows you to compare results from two different adjustment graphically.

If you only want to change the seasonal adjustment part and leave the pre-adjustment unchanged, you can plug-in the previous result to pre_adjustment, to speed up the computation.

```{r, AdjustmentX11, fig.width=15, fig.height=4.5, fig.align = 'center'}
result2 <-  dsa(series, 
                s7 = "x11",
                s365 = x11_method(sma = "S3x1", 
                                  sigma = c(2, 3)),
                xreg = regs,
                pre_processing = result)


compare_plot(result, result2)
```


## Adjustment with Seats
The adjustment called Seats is in fact just an Extended Airline based decomposition. Note that currently, multiplicative decompositions are not implemented in dsa().

```{r SeatsAdjustment, fig.width=15, fig.height=4.5, fig.align = 'center'}
result3 <- dsa(series, 
               s7 = seats_method(),
               s365 = seats_method(), 
               log = FALSE,
               xreg = regs)

compare_plot(result, result3)
```


## Day-of-the-month effects
The currency in circulation series analysed by Ollech (2018) included noticeable day-of-the-month effects. Not too many other daily time series include such effects, which is why by default, dsa2 does not include the estimation of these effects.

If you want to include it, you have to set the s31 parameter to a method of your choice


```{r DayOfMonth, fig.width=15, fig.height=4.5, fig.align = 'center'}
resultm <- dsa(series_withMonthlySeasonality, 
               s7 = NULL,
               s31 = seats_method(), 
               s365 = NULL,
               log = FALSE)

resultm2 <- dsa(series_withMonthlySeasonality, 
               s7 = NULL,
               s31 = x11_method(), 
               s365 = NULL,
               log = FALSE,
               pre_processing = resultm)

compare_plot(resultm, resultm2)
```


## References
Cleveland, R. B., W. S. Cleveland, J. E. McRae, and I. Terpenning. (1990). STL: A Seasonal-Trend Decomposition Procedure Based on Loess. Journal of Official Statistics, 6 (1), 3–73.

Ollech, D. (2018). Seasonal adjustment of daily time series. Bundesbank Discussion Paper 41/2018.

Ollech, D. (2021). Seasonal Adjustment of Daily Time Series. Journal of Time Series Econometrics, 13(2), 235-264.

Ollech, D. (2023). Economic analysis using higher-frequency time series: challenges for seasonal adjustment. Empirical Economics, 64(3), 1375-1398.