---
title: "BBkVignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BBkVignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

You can get Java 17 here: R:\Zentrale\Projekte\FB-S\Daten\DA\JDemetra\Anwendungen\VC\Java\JDK
Move and unpack the zip file to the right path: C:\\Workspace\\Java\\JDK\\ 
```{r start}
Sys.setenv("JAVA_HOME" = "C:\\Workspace\\Java\\JDK\\jdk-17.0.3+7") ## Always start with this
```


```{r setup}
library(dsa2)
```


```{r createSeries}
all <- tssim::sim_daily(N = 6, 
                        week_sd = 10,
                        year_sd = 10,
                        extra_smooth = TRUE,
                        calendar = list(which = "EasterMonday", 
                                        effect_size = 40, 
                                        from = -1, 
                                        to = 0))
series <- all$original


regs <- holidaysCentered[ ,c("EasterMonday", 
                            "EasterSunday")][seq.Date(from = start(series), 
                                                      length.out = length(series) + 365, 
                                                      by = "days")]
```



## Default adjustment

Currently, by default, STL is used for the seasonal adjustments
```{r Default, fig.width=7.5, fig.height=4.5}
result0 <- dsa(series)
result0
plot(result0)
# output(result) # Create HTML output
```

## Adjustment with missing values
```{r AdjustmentMissing, fig.width=7.5, fig.height=4.5}
seriesx <- series
seriesx[c(1, 499, 1010, 1748)] <- NA

resultx <- dsa(seriesx)
resultx
plot(resultx)
```


## Adjustment with regressors

```{r DefaultRegs, fig.width=7.5, fig.height=4.5}
result <- dsa(series,
              xreg = regs)
result
plot(result)
```

## Adjustment with X-11

Setting any of s7, s31 or s365 to "x11" uses the default adjustment of x11. Alternatively, x11_method() can be used, to specify details of the adjustment.
The function compare_plot() allows you to compare results from two different adjustment graphically.
Finally, if you only want to change the seasonal adjustment part and leave the pre-adjustment unchanged, you can plug-in the previous result to pre_adjustment, to speed up the computation.

```{r, AdjustmentX11, fig.width=7.5, fig.height=4.5}
result2 <-  dsa(series, 
                s7 = "x11", 
                s31 = "x11", 
                s365 = x11_method(sma = "S3x1", 
                                  sigma = c(2, 3)),
                xreg = regs,
                pre_processing = result)


compare_plot(result, result2)
```


## Adjustment with Seats
The adjustment called Seats is in fact just a Extended Airline based decomposition. Also, currently, multiplicative decompositions are not implemented.
```{r SeatsAdjustment, fig.width=7.5, fig.height=4.5}
result3 <- dsa(series, 
               s7 = seats_method(),
               s365 = seats_method(), 
               log = FALSE,
               xreg = regs)

compare_plot(result, result3)
```







